---
title: "HR Analytics"
author: "Alessio Crisafulli Carpani"
date: "2/11/2022"
output: html_document
---

```{r, include=FALSE, message=FALSE}
#Setup
setwd("~/Desktop/Study/Informatics/Projects/Ideas/HR Analytics")

pacman::p_load(tidyverse,reticulate)
```

```{python, include=FALSE, message=FALSE}
import pandas as pd
import matplotlib.pyplot as plt
```

# HR Analytics

## IBM Employee Attrition & Performance

The aim is to predict attrition of valuable employees and explore important questions such as 'show me a breakdown of distance from home by job role and attrition' or 'compare average monthly income by education and attrition'. This is a fictional data set created by IBM data scientists and it is possible to download it from [Kaggle](https://www.kaggle.com/pavansubhasht/ibm-hr-analytics-attrition-dataset){style="text-decoration: none;"}.

```{r}
IBM <- readr::read_csv("IBM-HR.csv", 
											 col_types = cols(
	Attrition = col_factor(levels = c("Yes", "No"))))

IBM
```

### Exploratory Data Analysis

```{python, include=FALSE}
a = r.IBM.describe().round(2)
for column in r.IBM.columns:
    print(f"{column}: Number of unique values {r.IBM[column].nunique()}")
```

For the purposes of this analysis the number of employees, the ID, the number of employees aged over 18 yo and standard hours, are not useful, therefore I am going to drop them.

```{python}
r.IBM.drop(['EmployeeCount', 'EmployeeNumber', 'Over18', 'StandardHours'], axis="columns", inplace=True)
```

#### Summary Statistics

Now, we can have a look to some summary statistics of the dataset. This is specially useful to understand the ranges of our ordinal variables:

```{r, message=FALSE}
reactable::reactable(py$a)
```

```{python, include=FALSE}
correlation = r.IBM.corr()

object_col = []
for column in r.IBM.columns:
	#Check variables that are categorical (objects in Python)
    if r.IBM[column].dtype == object and len(r.IBM[column].unique()) <= 30:
        object_col.append(column)
        print(f"{column} : {r.IBM[column].unique()}")
        print(r.IBM[column].value_counts())
        print("====================================")
```

#### Correlations

One of the first important tasks when doing exploratory data analysis is to see the correlation between our variables.

```{python}
import matplotlib.pyplot as plt
import seaborn as sns
plt.figure(figsize=(50, 50))
sns.set(font_scale=3)
sns.heatmap(r.IBM.corr(), annot=True, 
            linewidths=1.2,
            cmap=sns.color_palette("ch:s=.25,rot=-.25",                          as_cmap=True),
            annot_kws={"size": 25},
cbar_kws={"orientation": "horizontal"})
```

From the heatmap above we can asses some remarkable correlations:

-   Monthly income is highly correlated with Job level.

-   Job level is highly correlated with total working hours.

-   Monthly income is highly correlated with total working hours.

-   Age is also positively correlated with the Total working hours.

-   Marital status and stock option level are negatively correlated

> But what are the attrition rate?

```{python, include = FALSE}
#Number of observations for attrition attitude [YES,NO]
ratio = pd.DataFrame(
	((r.IBM.groupby(['Attrition'])['Attrition'].count()/len(r.IBM))).round(2)).T
```

```{r}
library(waffle)
waffle(c('Yes = 16%' = 16, 
				 'No = 84%' = 84), 
			 rows = 4, colors = c("orange", "lightgrey"),
			 size = 1,
  title = 'Attrition Rate (%)', legend_pos="bottom") + 
	theme(title = element_text(
		family = 'Arial',
		size = 5, face = 'bold'
	))
```
